Modelling Structured Societies: a Multi-relational Approach to Context Permeability

### 2015-06-20 doc/main.tex
Experimental Setup reviewed

	\section{Experimental Setup}
	\label{sec:experimental-setup}
\noindent In this section, we describe all the tools and processes necessary to produce the current research output. The experiments for our previously described simulation models were developed using the MASON~\cite{Luke2005} simulation framework written in Java. All the code for the simulation models can be found here \cite{Nunes:Software:11067}. Moreover, all the results presented in this article were made reproducible by using the \textit{statistical computing language R} \cite{R2008} and the R package \textit{knitr} \cite{knitr2014}. Knitr is used for generating reports that contain the code, the results of its execution (plots and tables), and the textual description for such results. All the R code used  to produce the data analysis and the configuration files used to reproduce the data can be found in \cite{NunesAntunes2014:Analysis:11898}.

%recap
For each configuration in our experiments, we performed 100 independent simulation runs. The results are thus analysed in terms of value distribution, average and standard deviation, or variance over the 100 runs. Like we described in section \ref{sec:sim_consensus}, our models are discrete-event models and in each cycle agents execute 
a step in a random order: this is so that this dynamical system --our simulation model-- is not influenced by the order in which agents are executed. Simulations end when total consensus is achieved or 2000 steps have passed.

@@

We use the method proposed in \cite{Barabasi1999} by Barab\'{a}si and Albert to construct the scale-free network instances using a preferential attachment. This model builds upon the perception of a common property of many large networks, a scale-free-power-law distribution of node connectivity. This feature was found to be a consequence of two generic mechanisms: networks expand continuously by the addition of new vertices, and new vertices attach preferentially to sites that are already well connected, being this mechanism more commonly known as ``preferential attachment''. In these networks, the probability $P(k)$ of two nodes being connected to each other decays as a power law, following $P(k) \sim k^{- \gamma}$.

The network instances were generated using the \textit{b-have network library} \cite{Nunes:Software:11069}. This is a Java library that allows for the manipulation of Network/Graph data structures. It also includes the more commonly used random network models from the literature.

@@

These properties are quite important to study social phenomena as they characterize structures that 
can be found in the real scenarios. The small-world networks generated by the models of Watts and Strogatz for instance \cite{Watts1998}, tries to model real-world social networks with short average path length, generate local clustering, and triadic closures between the nodes. One shortcoming of this model, is the fact that these networks do not possess node hubs found in many real-world social networks.

Since we construct scenarios where we overlap highly clustered k-regular networks and scale-free networks with node hubs, we are interested in what kind of properties emerge from merging these structures and how they are comparable to the existing models.

The networks produced by the \textit{b-have network library} \cite{Nunes:Software:11069} were exported to files and analysed using the \textit{igraph R package} \cite{igraph2006}.

>>>>

	\section{Experimental Setup}
	\label{sec:experimental-setup}
        \noindent In this section, we present all the tools and processes necessary to produce the
        current research output. The experiments were developed using the MASON~\cite{Luke2005}
        simulation framework written in Java. All the code for the simulation models can be found
        here \cite{Nunes:Software:11067}. Moreover, all the results presented in this article were
        made reproducible by using the \textit{statistical computing language R} \cite{R2008} and
        the R package \textit{knitr} \cite{knitr2014}. Knitr is used for generating reports that
        contain the code, the results of its execution (plots and tables), and the textual
        description for such results. All the R code used to produce the data analysis and the
        configuration files used to reproduce the data can be found in
        \cite{NunesAntunes2014:Analysis:11898}.

        % recap
        For each configuration in our experiments, we performed 100 independent simulation runs. The
        results are analysed in terms of value distribution, average and standard deviation, or
        variance over the 100 runs. Like we described in section \ref{sec:sim_consensus}, our models
        are discrete-event models. In each cycle agents execute a step in a random order: this is so
        that this dynamical system --our simulation model-- is not influenced by the order in which
        agents are executed. Simulations end when total consensus is achieved or 2000 steps have
        passed.

@@

        We use the method proposed in \cite{Barabasi1999} by Barab\'{a}si and Albert to construct
        the scale-free network instances using a preferential attachment. This model builds upon the
        perception of a common property of many large networks: a scale-free-power-law distribution
        of node connectivity. This feature was found to be a consequence of two generic mechanisms:
        networks expand continuously by the addition of new vertices, and new vertices attach
        preferentially to sites that are already well connected (more commonly known as
        ``preferential attachment''). In these networks, the probability $P(k)$ of two nodes being
        connected to each other decays as a power law, following $P(k) \sim k^{- \gamma}$.

        The network instances were generated using the \textit{b-have network library}
        \cite{Nunes:Software:11069}. This is a Java library that allows the creation and
        manipulation of network/graph data structures. It also includes the more commonly used
        random network models from the literature.

@@

        These properties characterise structures that can be found in the real scenarios. The
        small-world networks generated by the models of Watts and Strogatz for instance, model
        real-world social networks with short average path length, local clustering structures, and
        triadic closures between the nodes \cite{Watts1998}. One shortcoming of this model, is the
        fact that these networks do not possess node hubs found in many real-world social
        networks. (Nevertheless, these models are designed to create networks with specific
        properties, not as general models for all kinds of real social networks)

        Since we construct scenarios where we overlap highly clustered k-regular networks and
        scale-free networks with node hubs, we are interested in the kind of properties that emerge
        from merging these structures -- and if/how they are comparable to the existing models
        --. The networks were produced by the \textit{b-have network library}
        \cite{Nunes:Software:11069}, exported to files, and analysed using the \textit{igraph R
          package} \cite{igraph2006}.
